---
title: "Merging Datasets Using R"
author: "Assil NOUREDDINE"
email: "noureddine.assil@stud.hs-fresenius.de"
matriculation-number: "400924750"
date: "June 2025"
format: revealjs
bibliography: references.bib
csl: apa.csl
---

------------------------------------------------------------------------

------------------------------------------------------------------------

# Welcome

## Project Title

**Merging Datasets Using R**

### Assil NOUREDDINE

Matriculation Number: 400924750\
Email: noureddine.assil\@stud.hs-fresenius.de

------------------------------------------------------------------------

# Abstract

This presentation explains how to merge datasets in R using `tidyverse`. We use fictional employee and salary datasets to demonstrate `left_join()` and `inner_join()`, and visualize the results.

------------------------------------------------------------------------

# Motivation

-   Merging datasets is essential in data science
-   Data often comes from multiple sources
-   We need clean, complete data to analyze

------------------------------------------------------------------------

# The Datasets

-   `employees.csv`\
-   `salaries.csv`\
    Both use `employee_id` as the linking key.

------------------------------------------------------------------------

# Load and Preview Data

```{r}
library(tidyverse)
employees <- read_csv("data/employees.csv")
salaries <- read_csv("data/salaries.csv")

glimpse(employees)
glimpse(salaries)
```

------------------------------------------------------------------------

# Left Join

```{r}
left_joined <- left_join(employees, salaries, by = "employee_id")
glimpse(left_joined)
```

------------------------------------------------------------------------

# Inner Join

```{r}
inner_joined <- inner_join(employees, salaries, by = "employee_id")
```

------------------------------------------------------------------------

# Analysis by Department

```{r}
left_joined %>%
  group_by(department) %>%
  summarise(
    avg_salary = mean(salary, na.rm = TRUE),
    count = n()
  )
```

------------------------------------------------------------------------

# Visualization

```{r}
left_joined %>%
  group_by(department) %>%
  summarise(avg_salary = mean(salary, na.rm = TRUE)) %>%
  ggplot(aes(x = department, y = avg_salary)) +
  geom_col(fill = "steelblue") +
  theme_minimal()
```

------------------------------------------------------------------------

# Discussion

-   Differences in salary between departments
-   Why left join was useful
-   Potential causes: experience, role, etc.

------------------------------------------------------------------------

# Conclusion

-   Merged datasets using `dplyr` joins
-   Showed how to summarize and plot
-   Useful for real-world reporting

------------------------------------------------------------------------

# References

Add citations from your `references.bib` here using `@key`.

------------------------------------------------------------------------

# Thank You

Any questions?
